GVN-Hoist: Hoisting Computations from Branches

### 2016-08-11 hoist.tex
restructure

\section{Code hoisting policies}
The amount of hoisting depends on whether we collect GVN of instructions
on-demand (pessimistic) or, before finding candidates (optimistic).It also
depends on the generality of the GVN algorithm, however, that analysis is beyond
the scope of this paper.

@@

\begin{verbatim}
BasicBlock B1, B2, B3 // B1 != B2 != B3
DFSIn(B1) = depth first discovery time stamp
DFSIn(B1,B2) = |DFSIn(B1) - DFSIn(B2)|
DFSIn(B1,B3) = |DFSIn(B1) - DFSIn(B3)|

@@

In the presence of fully redundant computations in the list of hoistable
candidates this equation may not hold. For example, if there is another
instruction I4 which is dominated by I1, but I4 is present in a basic block
which is farther than DFSIn(I1, I2) i.e., $DFSIn(I1,I2) < DFSIn(I1,I4)$;
however, $BBDist(I1,I4) = 0 < BBDist(I1,I2) >=1$

So by sorting candidates w.r.t. their DFSIn numbers:
\begin{enumerate}
\item would make fewer checks for legality and profitability.
\item hoisting point will be closer so, the intersection of live-range of the
  instruction with other instructions will be minimal.
\end{enumerate}

In the case where all the candidates cannot be hoisted at a common hoist point
because of:
\begin{enumerate}
\item Safety conditions are not met.
\item Live range would increase too much.
\end{enumerate}

In our current implementation we keep as many candidates in one set as possible
unless (greedy approach). We split the list at a point where the legality checks
fail to hoist subset of candidates which are legal to hoist and, start finding
new hoisting point for the remaining ones.

TODO: Put an example to illustrate the point.

In the case when more than one computation from the same basic block
shows up as redundant.

@@

\subsubsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that could
be hoisted in the program, matching the complexity of PRE on SSA form.  The
analysis phase is based on the Global Value Numbering (GVN), the same analysis
used for PRE, followed by the computation of a partition of identical
expressions to be hoisted in a same location to guarantee safety properties and
program performance, and followed by a simple code generation that adds the
identified instruction in the hoisting point and removes all the now redundant
expressions.

>>>>

\subsection{Illustrative Example}
In the case when more than one computation from the same basic block
shows up as redundant.

\section{Code hoisting policies}
The amount of hoisting depends on whether we collect GVN of instructions
before finding candidates (optimistic) or, on-demand (pessimistic). It also
depends on the generality of the GVN algorithm, however, that analysis is beyond
the scope of this paper.

@@

\begin{verbatim}
// B1 != B2 != B3
BasicBlock B1, B2, B3
DFSIn(B1) = depth first discovery time stamp
DFSIn(B1,B2) = |DFSIn(B1) - DFSIn(B2)|
DFSIn(B1,B3) = |DFSIn(B1) - DFSIn(B3)|

@@

In the presence of fully redundant computations in the list of hoistable
candidates this equation may not hold. For example, if there is another
instruction I4 which is dominated by I1, but I4 is present in a basic block
which is farther than DFSIn(I1, I2) i.e., $DFSIn(I1,I2) < DFSIn(I1,I4)$;
however, $BBDist(I1,I4) = 0 < BBDist(I1,I2) >=1$.

So by sorting candidates w.r.t. their DFSIn numbers:
\begin{enumerate}
\item would make fewer checks for legality and profitability.
\item hoisting point will be closer, so the intersection of live-range of the
  instruction with other instructions will be minimal.
\end{enumerate}

In our current implementation we keep as many candidates in one set as possible
(greedy approach). We split the list at a point where the legality checks fail
to hoist subset of candidates which are legal to hoist and then start finding
new hoisting point for the remaining ones. So there are two limitations of the
current implementation: first, sorting by DFSIn numbers does not give the right
order when there are fully redundant instructions, second, it lacks cost model
for partitioning in order to preserve/improve performance.

@@

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that could
be hoisted in the program, matching the complexity of PRE on SSA form.  The
analysis phase is based on the Global Value Numbering (GVN), the same analysis
used for PRE, followed by the computation of a partition of identical
expressions to be hoisted in a same location to guarantee safety properties and
program performance, and followed by a simple code generation that adds the
identified instruction in the hoisting point and removes all the now redundant
expressions.
